### 2025-05-05 paper/ecai2025/sects/discussion.tex
minor on fig 1

\section{Discussion}
In the following, we discuss the results obtained in \Cref{sect:experiments} and provide some insights into the performance of \fname{} compared to SHAC and PPO trying to answer the research questions posed in \Cref{sect:introduction}. Note that \fname{} does not aim to outperform SHAC, it merely aims to perform comparably with SHAC while being applicable even in non-differentiable environment. 
\begin{rqbox}
\textbf{Can we train a neural network to approximate the gradients of a differentiable simulator?}
\end{rqbox}

The proposed framework, \fname{}, appears to be successful in emulating the behavior of SHAC in differentiable environments such as transport and sampling. Therefore, we can conclude that it is definitely possible to approximate the gradients of a differentiable environment with a neural network. However, we note that this approach, while obtaining comparable results in most setups, it tends to display higher variance. Both neural networks representing the reward and transition functions require sufficient training before becoming effective, which may entail multiple episodes. Interestingly, \fname{} outperforms SHAC in the Sampling scenario. A possible explanation is that the complex dynamics of the environment may lead to unstable gradients.

@@

\begin{rqbox}
\textbf{How does our algorithm compare to PPO and SHAC in both single-agent and multi-agent settings?}
\end{rqbox}

The proposed algorithm appears capable of outperforming PPO across all scenarios showing also a better sample efficiency.

@@

\begin{rqbox}
\textbf{How does the performance of these algorithms change as the search space increases?}
\end{rqbox}

To analyze the impact of increasing search spaces, we primarily focused on scenarios with varying numbers of agents (up to 5 in the main experiments). The results demonstrate that performance patterns differ across scenarios. In the Dispersion scenario, \fname{} shows increasing difficulty in finding optimal policies as the number of agents grows. Conversely, in Transport and Sampling scenarios, the system appears to converge more quickly with more agents. This counter intuitive behavior likely stems from the fact that, despite larger search spaces, the fundamental tasks in Sampling and Transport become relatively easier to solve with multiple agents, making policy discovery more straightforward. Even with approximated world dynamics (transition and reward functions), the guidance provided is sufficient to lead policies toward optimal behavior.

>>>>

In the following, we discuss the results obtained in \Cref{sect:experiments} and provide some insights into the performance of \fname{} compared to SHAC and PPO, addressing the research questions posed in \Cref{sect:introduction}. Note that \fname{} does not aim to outperform SHAC, but rather to perform comparably while being applicable even in non-differentiable environments.

\subsection{Gradient Approximation with Neural Networks}
\begin{rqbox}
\textbf{Can we train a neural network to approximate the gradients of a differentiable simulator?}
\end{rqbox}

The proposed framework, \fname{}, appears to be successful in emulating the behavior of SHAC in differentiable environments such as transport and sampling. Therefore, we can conclude that it is definitely possible to approximate the gradients of a differentiable environment with a neural network. However, we note that this approach, while obtaining comparable results in most setups, tends to display higher variance. Both neural networks representing the reward and transition functions require sufficient training before becoming effective, which may entail multiple episodes. Interestingly, \fname{} outperforms SHAC in the Sampling scenario. A possible explanation is that the complex dynamics of the environment may lead to unstable gradients.

@@

\subsection{Comparison to PPO and SHAC}
\begin{rqbox}
\textbf{How does our algorithm compare to PPO and SHAC in both single-agent and multi-agent settings?}
\end{rqbox}

The proposed algorithm appears capable of outperforming PPO across all scenarios, also showing better sample efficiency.

@@

\subsection{Scaling with Search Space}
\begin{rqbox}
\textbf{How does the performance of these algorithms change as the search space increases?}
\end{rqbox}

To analyze the impact of increasing search spaces, we primarily focused on scenarios with varying numbers of agents (up to 5 in the main experiments). The results demonstrate that performance patterns differ across scenarios. In the Dispersion scenario, \fname{} shows increasing difficulty in finding optimal policies as the number of agents grows. Conversely, in Transport and Sampling scenarios, the system appears to converge more quickly with more agents. This counterintuitive behavior likely stems from the fact that, despite larger search spaces, the fundamental tasks in Sampling and Transport become relatively easier to solve with multiple agents, making policy discovery more straightforward. Even with approximated world dynamics (transition and reward functions), the guidance provided is sufficient to lead policies toward optimal behavior.

### 2025-05-05 paper/ecai2025/sects/shacpp.tex
minor on fig 1

\begin{figure*}[t]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figs/shac.pdf}
        \caption{SHAC}%
        \label{fig:shac}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figs/shacpp.pdf}
        \caption{\fname{}}%
        \label{fig:shacpp}
    \end{subfigure}
    \caption{
        Comparison between SHAC and \fname{} during a rollout of 4 steps.
        Nodes with labels (such as $r_i$ or $r'_i$) values. Nodes without labels (such as $\pi$ or $R$) represent functions. Dashed lines are used to represent non differentiability. $a_i$ and $r_i$ represent the collection of actions and rewards for all agents}\label{fig:shac-shacpp}
\end{figure*}

>>>>

\begin{figure*}[t]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figs/shac.pdf}
        \caption{SHAC}%
        \label{fig:shac}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figs/shacpp.pdf}
        \caption{\fname{}}%
        \label{fig:shacpp}
    \end{subfigure}
    \caption{
        Comparison between SHAC and \fname{} during a rollout of 4 steps.
        Nodes with labels (such as $r_i$ or $r'_i$) indicate values, while nodes without labels (such as $\pi$ or $R$) represent functions. Dashed lines denote non-differentiability. $a_i$ and $r_i$ refer to the collection of actions and rewards for all agents. Note that the main advantage of \fname{} is that the transition function $F$ on the right can be non-differentiable, unlike the left $F$ in SHAC, which must be differentiable.
        }\label{fig:shac-shacpp}
\end{figure*}
